Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-25 cav2017/cav2017.tex
write initial version of the conclusion

\begin{abstract}
%
We present a sound and automated approach to synthesize safe digital
feedback controllers for physical plants represented as linear, time
invariant models.  Models are described as dynamical equations with inputs,
evolving over a continuous state space and accounting for errors due to the
digitalization of signals by the controller.  The approach works in two
stages, leveraging counterexample guided inductive synthesis (CEGIS) and
reachability analysis.  CEGIS synthesises a static feedback controller that
stabilizes the system under some restrictions given by the safety of the
reach space.  Safety is verified either via BMC or abstract acceleration; if
the verification step fails, we generate an alternative controller by
generalizing the counterexample.  We synthesize stable and safe controllers
for intricate physical plant models from the digital control literature.
%
\keywords{
State-space dynamical models of physical systems; 
Digital controllers; 
Analogue-to-digital converters; 
Time sampling; 
quantization; 
Fixed-point arithmetic; 
CEGIS; 
safety requirements. 
}
\end{abstract}

@@

In this work, we propose the synthesis of \emph{safe} control algorithms for
LTI models taking into consideration the continuous dynamics of the plant, 
alongside the discrete dynamics of the controller and the effects of its 
implementation, as well as the hybrid elements that relate them.  
%% In this paper, we synthesize controllers for hybrid
%% closed-loop systems, i.e., software-implemented embedded controllers along
%% with a model of their physical environment (the plant), that are
%% \emph{safe}.
Due to the complexity of such systems, in this work we focus 
on linear models with known  implementation features (e.g., number of bits, fixed-point
arithmetic). We specify the safety requirement over the state space. This is a property often 
overlooked in feedback control synthesis, but which plays an important role 
in real-world systems.

@@

{\em The first approach} uses a {\em naive}
technique that starts by devising a digital controller that stabilizes the system
whilst remaining safe for a 
pre-selected time horizon and a single initial state; then, it verifies its potential safety by unfolding
the dynamics of the system, considering the full hyper-cube of initial states, and checking for a {\em completeness threshold}, i.e.,
the number of iterations required to sufficiently unwind
the closed-loop state-space system such
that the boundaries are not violated for any larger number of iterations.
As it requires explicit unfolding up to the completeness
threshold, this approach may suffer from state space explosion.

@@

%-------------------------------
\subsection*{Related work}
%\label{sec:relw}
%-------------------------------

@@

The tool Pessoa~\cite{mazo2010pessoa} synthesizes
correct-by-design embedded control software, in a Matlab ToolBox.  It is based on the abstraction
of a physical system to an equivalent finite-state machine and on the
computation of reachability properties over the finite-state machine.  Based
on this safety specification over state values, Pessoa can synthesize
embedded controller software for a range of properties.
The embedded controller software can be more complicated than
the state feedback control we synthesised, and the properties 
available cover more detail.
%based on a safety specification over state values, as well as specification on characteristic functions of the controller. 
However, Pessoa does not have native inclusion of the quantization error
introduced by Digital-to-Analogue/Analogue-to-Digital conversion needed for
synthesizing fixed-point digital controllers.

@@

We illustrate our approach with an example,
extracted from the literature~\cite{Franklin15}.
%
We start with a candidate solution with all coefficients zero, $K=[0
  \ 0 \ 0]^T$, and a precision of $I_p=13$, $F_p=3$.  In the first
{\sc verify} stage, the {\sc safety} check finds the counterexample
%
$ x_0 = [-0.5 \ 0.5 \ 0.5] $.
%
After adding the new counterexample to its sets of {\sc inputs}, {\sc
  synthesize} finds the candidate solution $K=[0 \ 0
  \ 0.00048828125]^T$, which prompts the {\sc safety} verifier to
return $x_0=[-0.5 -0.5 -0.5]$ as the new counterexample.

In the subsequent iteration, the synthesizer is unable to find further 
suitable candidates and it returns UNSAT, meaning that the current precision is
insufficient.  Consequently, we increase the precision to $I_p=17$,
$F_p=7$.
%
Since the previous counterexamples were obtained at lower precision,
we remove them from the set of counterexamples.  Back in the {\sc
  synthesize} phase, we re-start the process with a candidate solution
with all coefficients $0$.  Next, the {\sc safety} verification stage
provides the first counterexample at higher precision, $x_0=[-0.5
  \ 0.5 \ 0.5]$ and {\sc synthesize} finds $K=[0 \ 0.01171875
  \ 0.015625]^T$ as a candidate that eliminates this counterexample.
However, this candidate triggers the counterexample
$x_0=[0.5\ -0.5\ -0.5]$ found again by the {\sc safety} verification
stage.  In the next iteration, we get the candidate $K=[0 \ 0
  \ -0.015625]$, followed by the counterexample $x_0 = [0.5 \ 0.5
  \ 0.5]$. Finally, {\sc synthesize} finds the candidate $K=[0.01171875
  \ -0.013671875 \ -0.013671875]^T$, which is validated as a final
solution by all verification stages.

@@

\end{tabular}
\caption{Experimental results with runtime in milliseconds. \label{tab:results}}
\end{table}

@@

>>>>

\begin{abstract}
%
We present a sound and automated approach to synthesize safe digital
feedback controllers for physical plants represented as linear, time
invariant models.  Models are described as dynamical equations with inputs,
evolving over a continuous state space and accounting for errors due to the
digitalization of signals by the controller.  The approach works in two
stages, leveraging counterexample guided inductive synthesis (CEGIS) and
reachability analysis.  CEGIS synthesises a static feedback controller that
stabilizes the system under some restrictions given by the safety of the
reach space.  Safety is verfied iether via BMC or abstract acceleration; if
the verification step fails, we generate an alternative controller by
generalizing the counterexample.  We synthesize stable and safe controllers
for intricate physical plant models from the digital control literature.
%
\keywords{
State-space dynamical models of physical systems; 
Digital controllers; 
Analogue-to-digital converters; 
Time sampling; 
quantization; 
Fixed-point arithmetic; 
CEGIS; 
safety requirements. 
}
\end{abstract}

@@

In this work, we propose the synthesis of \emph{safe} control algorithms for
LTI models taking into consideration the continuous dynamics of the plant, 
alongside the discrete dynamics of the controller and the effects of its 
implementation, as well as the hybrid elements that relate them.  
%% In this paper, we synthesize controllers for hybrid
%% closed-loop systems, i.e., software-implemented embedded controllers along
%% with a model of their physical environment (the plant), that are
%% \emph{safe}.
Due to the complexity of such systems, in this work we focus 
on linear models with known  implementation features (e.g., number of bits, fixed-point
arithmetic), and perform synthesis of safe digital controllers, a property often 
overlooked in feedback control synthesis, but which plays an important role 
in real-world systems.

@@

{\em The first approach} uses a {\em naive}
technique that starts by devising a digital controller that stabilizes the system
whilst remaining safe for a 
pre-selected time horizon; then, it verifies its potential safety by unfolding
the dynamics of the system and checking for a {\em completeness threshold}, i.e.,
the number of iterations required to sufficiently unwind
the closed-loop state-space system such
that the boundaries are not violated for any larger number of iterations.
As it requires explicit unfolding up to the completeness
threshold, this approach may suffer from state space explosion.

@@

%-------------------------------
\section{Related work}
\label{sec:relw}
%-------------------------------

@@

The tool Pessoa~\cite{mazo2010pessoa} synthesizes
correct-by-design embedded control software, in a Matlab ToolBox.  It is based on the description
of a physical system as an equivalent finite-state machine and on the
computation of reachability properties over the finite-state machine.  Based
on this safety specification over state values, Pessoa can synthesize
embedded controller software for a range of properties.
The embedded controller software can be more complicated than
the state feedback control we synthesised, and the properties 
available cover more detail.
%based on a safety specification over state values, as well as specification on characteristic functions of the controller. 
However, Pessoa does not have native inclusion of the quantization error
introduced by Digital-to-Analogue/Analogue-to-Digital conversion needed for
synthesizing fixed-point digital controllers.

@@

We illustrate our approach with \todo{what example is this?} an example,
extracted from the literature~\cite{Franklin15}.
%
We start with a candidate solution with all coefficients zero, $K=[0
  \ 0 \ 0]^T$, and a precision of $I_p=13$, $F_p=3$.  In the first
{\sc verify} stage, the {\sc safety} check finds the counterexample
%
$ x_0 = [-0.5 \ 0.5 \ 0.5] $.
%
After adding the new counterexample to its sets of {\sc inputs}, {\sc
  synthesize} finds the candidate solution $K=[0 \ 0
  \ 0.00048828125]^T$, which prompts the {\sc safety} verifier to
return $x_0=[-0.5 -0.5 -0.5]$ as the new counterexample.

In the subsequent iteration, the synthesizer is unable to find further 
suitbale candidates and it returns UNSAT, meaning that the current precision is
insufficient.  Consequently, we increase the precision to $I_p=17$,
$F_p=7$.
%
Since the previous counterexamples were obtained at lower precision,
we remove them from the set of counterexamples.  Back in the {\sc
  synthesize} phase, we re-start the process with a candidate solution
with all coefficients $0$.  Next, the {\sc safety} verification stage
provides the first counterexample at higher precision, $x_0=[-0.5
  \ 0.5 \ 0.5]$ and {\sc synthesize} finds $K=[0 \ 0.01171875
  \ 0.015625]^T$ as a candidate that eliminates this counterexample.
However, this candidate triggers the counterexample
$x_0=[0.5\ -0.5\ -0.5]$ found again by the {\sc safety} verification
stage.  In the next iteration, we get the candidate $K=[0 \ 0
  \ -0.015625]$, followed by the counterexample $x_0 = [0.5 \ 0.5
  \ 0.5]$. Finally, {\sc synthesize} finds the candidate $K=[0.01171875
  \ -0.013671875 \ -0.013671875]^T$, which is validated as a final
solution by all verification stages.

@@

\end{tabular}
\caption{Experimental results with runtime in milliseconds \label{tab:results}}
\end{table}

@@

We have presented two approaches to synthesize state feedback controllers 
that ensure both stability and safety over the state-space representation.
The first approach relies on explicit unfolding of the system's dynamics 
up to the completeness threshold, whilst the second one applies abstraction 
refinement and acceleration to increase speed whilst retaining soundness.
Both approaches are novel in the control system literature in that it puts
forward a fully automated synthesis approach that is algorithmically and
numerically sound, considering the various error sources in the implementation 
of the control algorithm and in the computational modeling of plant dynamics.
Our experimental results show that both approaches are able
to synthesize stable controllers in most benchmarks within a reasonable
amount of time fully automatically.  In particular, both approaches
complement each other in the benchmark coverage and together 
solve all benchmarks provided in the set.
Future work will be the extension of these approaches to consider performance 
requirements while synthesizing digital controllers.
